### 2019-07-09 2-Chapters/4-Chapter/CrownCom_17.git/IoT_slotted.tex
Corrections fautes orthographes

\textbf{One gateway and many devices.}
%
We consider the system model presented in Figure~\ref{fig:41:system_model1}, where a set of devices all sends uplink packets to a (unique) network gateway.
Messages can be sent in a fixed number $K\geq2$ of wireless channels,
that are assumed to be orthogonal, that is, a message in channel $k$ will not cause collision to a message in another channel $j \neq k$.
% (\eg, $K=4$ or $K=16$ for instance), which are centered indifferently around the $433.5\;\mathrm{MHz}$ or the $877\;\mathrm{MHz}$ ISM bands (as it is chosen for the implementation presented in Section~\ref{sec:4:gnuradio}).
The communication between IoT devices and this gateway is done through a simple pure acknowledged ALOHA-based protocol where devices transmit uplink packets of fixed duration whenever they want.
We mean here by ALOHA-based that \emph{devices do not use sensing}, and \emph{we do not consider retransmissions of packets} (we relax this hypothesis in Section~\ref{sec:4:retransmissions}).
%
% Note that this choice of number of channels and of frequency bands is for instance the one of the current LoRa standard in Europe, but of course the precise choice does not really matter, and future standards can choose different parameters.

@@

The devices can transmit their packets in one of the $K$ channels. In the case where the gateway --of the corresponding IoT network-- receives an uplink message in one channel, it transmits an acknowledgement to the end-device in the same channel, after a fixed delay.
This is realistic in IoT networks such as networks based on LoRaWAN.
% (of $1\;\mathrm{s}$ in our implementation).
In a more general setting, if there were more than one gateway in the same IoT network being considered, the network would reply to the device by letting only one gateway sends back an acknowledgement.
The natural choice is for the network to send this acknowledgement by the gateway which received the uplink message.
For simplicity but without loss of generality, we consider a network with only one gateway,
but still it is important to observe the distinction between the gateway, which is simply a RF node, and the IoT network in charge of receiving, handling, and replying to the incoming uplink messages from the IoT devices being paired in the network.
%

These communications operate in unlicensed ISM bands and, consequently, they can suffer from interference generated by uncoordinated co-localized and/or neighboring networks (two are shown in \textcolor{orange}{orange} in Figure~\ref{fig:41:system_model1}, in left or right).
This interfering traffic is uncontrolled, and is most likely unevenly distributed over the $K$ different channels,
as each IoT protocol or IoT network can choose to use a different subset of channels.
%
In order to simulate networks designed for the IoT, we consider a protocol \textbf{with no sensing} and no repetition of uplink messages. The gateway is in charge of sending back an acknowledgement, after some fixed-time delay, to any device paired with this gateway in this network (\ie, in \textcolor{blue}{blue} in Figure~\ref{fig:41:system_model1}) which succeeded in sending an uplink packet.
%
By considering a small number of orthogonal wireless channels, and a unique PHY layer configuration (\ie, modulation, waveform, etc), and in case of a non-uniform traffic in the different channels,
the device can improve their usage of the network if they are able to \emph{learn} on the fly the best channels to use, that is, the most vacant one. Indeed, in this model, the \emph{quality} of the channels (\ie, the ressources) are identified by their vacancy, or one minus their occupancy rates.

@@

\begin{figure}[!h]
    \centering
    \includegraphics[scale=0.40]{protocol.eps}
    \caption[The considered time-frequency slotted protocol. Each frame is composed by a fixed duration uplink slot in which the end-devices transmit their (uplink) packets. If a packet is well received, the gateway replies by transmitting an \Ack, after the ack delay.]{The considered time-frequency ALOHA-based communication protocol \cite{Abramson1970,Roberts75}. Each frame is composed by a fixed duration \textcolor{blue}{uplink slot} in which the end-devices transmit their \textcolor{blue}{(uplink) packets}. If a packet is well received, the gateway replies by transmitting an \textcolor{darkgreen}{\Ack}, after the ack delay.}
    \label{fig:41:protocol}
\end{figure}

\paragraph{Slotted protocol.}
%
For simplicity, and as in all this thesis, we only consider a discrete protocol (\ie, slotted).
As illustrated in Figure~\ref{fig:41:protocol}, we suppose a slotted protocol, in both time and frequency.
All devices share a synchronized time, and know in advance $K$, the finite number of available RF channels.
In each time slot, the devices try to send packets to the unique gateway, which listens continuously to all channels, following an ALOHA-based communication protocol \cite{Abramson1970,Roberts75}, with no sensing.
Each time slot is divided in two parts: first for uplink communications in which data packets are sent by end-devices to the gateway.
In one channel, if only one packet is sent in this part of the slot, the gateway can decode it and sends an acknowledgement to the device in the second part (on the same channel).
If two or more devices send an uplink packet in the same slot, the uplink packets collide, and the acknowledgement (\Ack) is not transmitted. In this case, we say that there is a \emph{collision} in this time slot.
In other words, if the gateway cannot successfully decode the incoming (uplink) messages because they were corrupted by collisions, it does not send any \Ack{} back.
This way, no collision can occur on the downlink messages, easing the analysis of collisions.

@@

We can see in Figure~\ref{fig:41:from10to100} that the TS algorithm (\textcolor{red}{in red $+$}) outperforms the \UCB{} algorithm (\textcolor{blue}{in blue} $*$), when the number of end-devices is below 50\%. When the number of end-devices is higher, both algorithms have almost the same performance, and perform well after a small number of transmissions (\ie, they show quick convergence).
Moreover, we can see in Figures~\ref{fig:41:10intelligent}, \ref{fig:41:30intelligent}, and \ref{fig:41:50intelligent} that both have better success rate than the random policy and the probability of successful transmission is between the optimal oracle and suboptimal oracle policies.
For instance, for $10\%$ of dynamic devices, after about $1000$ transmissions, using \UCB{} over the naive uniform policy improved the successful transmission rate from $83\%$ to $88\%$, and using Thompson Sampling improved it to $89\%$.
Increasing the number of end-devices decreases the gap between the optimal and random policies:
as expected intuitively, the more dynamic devices, the less useful are learning algorithms, and basically for networks with only dynamic devices, the random policy is as efficient as the optimal one, as seen in Figures~\ref{fig:41:100intelligent} and on the right end side of Figure~\ref{fig:41:perf_learning}.

>>>>

\textbf{One gateway and many devices.}
%
We consider the system model presented in Figure~\ref{fig:41:system_model1}, where a set of devices all sends up-link packets to a (unique) network gateway.
Messages can be sent in a fixed number $K\geq2$ of wireless channels,
that are assumed to be orthogonal, that is, a message in channel $k$ will not cause collision to a message in another channel $j \neq k$.
% (\eg, $K=4$ or $K=16$ for instance), which are centered indifferently around the $433.5\;\mathrm{MHz}$ or the $877\;\mathrm{MHz}$ ISM bands (as it is chosen for the implementation presented in Section~\ref{sec:4:gnuradio}).
The communication between IoT devices and this gateway is done through a simple pure acknowledged ALOHA-based protocol where devices transmit up-link packets of fixed duration whenever they want.
We mean here by ALOHA-based that \emph{devices do not use sensing}, and \emph{we do not consider retransmissions of packets} (we relax this hypothesis in Section~\ref{sec:4:retransmissions}).
%
% Note that this choice of number of channels and of frequency bands is for instance the one of the current LoRa standard in Europe, but of course the precise choice does not really matter, and future standards can choose different parameters.

@@

The devices can transmit their packets in one of the $K$ channels. In the case where the gateway --of the corresponding IoT network-- receives an up-link message in one channel, it transmits an acknowledgement to the end-device in the same channel, after a fixed delay.
This is realistic in IoT networks such as networks based on LoRaWAN.
% (of $1\;\mathrm{s}$ in our implementation).
In a more general setting, if there were more than one gateway in the same IoT network being considered, the network would reply to the device by letting only one gateway sends back an acknowledgement.
The natural choice is for the network to send this acknowledgement by the gateway which received the up-link message.
For simplicity but without loss of generality, we consider a network with only one gateway,
but still it is important to observe the distinction between the gateway, which is simply a RF node, and the IoT network in charge of receiving, handling, and replying to the incoming up-link messages from the IoT devices being paired in the network.
%

These communications operate in unlicensed ISM bands and, consequently, they can suffer from interference generated by uncoordinated co-localized and/or neighboring networks (two are shown in \textcolor{orange}{orange} in Figure~\ref{fig:41:system_model1}, in left or right).
This interfering traffic is uncontrolled, and is most likely unevenly distributed over the $K$ different channels,
as each IoT protocol or IoT network can choose to use a different subset of channels.
%
In order to simulate networks designed for the IoT, we consider a protocol \textbf{with no sensing} and no repetition of up-link messages. The gateway is in charge of sending back an acknowledgement, after some fixed-time delay, to any device paired with this gateway in this network (\ie, in \textcolor{blue}{blue} in Figure~\ref{fig:41:system_model1}) which succeeded in sending an up-link packet.
%
By considering a small number of orthogonal wireless channels, and a unique PHY layer configuration (\ie, modulation, waveform, etc), and in case of a non-uniform traffic in the different channels,
the device can improve their usage of the network if they are able to \emph{learn} on the fly the best channels to use, that is, the most vacant one. Indeed, in this model, the \emph{quality} of the channels (\ie, the resources) are identified by their vacancy, or one minus their occupancy rates.

@@

\begin{figure}[!h]
    \centering
    \includegraphics[scale=0.40]{protocol.eps}
    \caption[The considered time-frequency slotted protocol. Each frame is composed by a fixed duration up-link slot in which the end-devices transmit their (up-link) packets. If a packet is well received, the gateway replies by transmitting an \Ack, after the ack delay.]{The considered time-frequency ALOHA-based communication protocol \cite{Abramson1970,Roberts75}. Each frame is composed by a fixed duration \textcolor{blue}{up-link slot} in which the end-devices transmit their \textcolor{blue}{(up-link) packets}. If a packet is well received, the gateway replies by transmitting an \textcolor{darkgreen}{\Ack}, after the ack delay.}
    \label{fig:41:protocol}
\end{figure}

\paragraph{Slotted protocol.}
%
For simplicity, and as in all this thesis, we only consider a discrete protocol (\ie, slotted).
As illustrated in Figure~\ref{fig:41:protocol}, we suppose a slotted protocol, in both time and frequency.
All devices share a synchronized time, and know in advance $K$, the finite number of available RF channels.
In each time slot, the devices try to send packets to the unique gateway, which listens continuously to all channels, following an ALOHA-based communication protocol \cite{Abramson1970,Roberts75}, with no sensing.
Each time slot is divided in two parts: first for up-link communications in which data packets are sent by end-devices to the gateway.
In one channel, if only one packet is sent in this part of the slot, the gateway can decode it and sends an acknowledgement to the device in the second part (on the same channel).
If two or more devices send an up-link packet in the same slot, the up-link packets collide, and the acknowledgement (\Ack) is not transmitted. In this case, we say that there is a \emph{collision} in this time slot.
In other words, if the gateway cannot successfully decode the incoming (up-link) messages because they were corrupted by collisions, it does not send any \Ack{} back.
This way, no collision can occur on the down-link messages, easing the analysis of collisions.

@@

We can see in Figure~\ref{fig:41:from10to100} that the TS algorithm (\textcolor{red}{in red $+$}) outperforms the \UCB{} algorithm (\textcolor{blue}{in blue} $*$), when the number of end-devices is below 50\%. When the number of end-devices is higher, both algorithms have almost the same performance, and perform well after a small number of transmissions (\ie, they show quick convergence).
Moreover, we can see in Figures~\ref{fig:41:10intelligent}, \ref{fig:41:30intelligent}, and \ref{fig:41:50intelligent} that both have better success rate than the random policy and the probability of successful transmission is between the optimal oracle and sub-optimal oracle policies.
For instance, for $10\%$ of dynamic devices, after about $1000$ transmissions, using \UCB{} over the naive uniform policy improved the successful transmission rate from $83\%$ to $88\%$, and using Thompson Sampling improved it to $89\%$.
Increasing the number of end-devices decreases the gap between the optimal and random policies:
as expected intuitively, the more dynamic devices, the less useful are learning algorithms, and basically for networks with only dynamic devices, the random policy is as efficient as the optimal one, as seen in Figures~\ref{fig:41:100intelligent} and on the right end side of Figure~\ref{fig:41:perf_learning}.

### 2019-07-09 2-Chapters/4-Chapter/IEEE_WCNC_2019__DemoICT.git/main.tex
Corrections fautes orthographes

We describe the way we implemented a demo where we evaluate MAB algorithms, used in combination with a pure ALOHA-based protocol, such as the ones employed in LPWAN.
This demonstration is the first implementation which aims at assessing the potential gain of MAB learning algorithms in IoT scenarios.
%
We use a TestBed designed in 2017 by the SCEE team at the Rennes campus of CentraleSupélec \cite[Appendix~3]{Bodinier17}, containing different USRP boards \cite{USRPDocumentation}, controlled by a single laptop running the GNU Radio software \cite{GNURadioDocumentation},
and where the intelligence of each device corresponds to a learning algorithm, implemented as a GNU Radio block \cite{GNURadioCompanionDocumentation} and written in Python or \texttt{C++}.

@@

Regarding the physical layer, we consider a QPSK constellation (\emph{Quadrature Phase-Shift Keying}, \eg, see references in \cite{Bodinier17}). Moreover, we use simplified packets composed of two parts.
The first part is the \emph{preamble} which is used for the purpose of synchronization (phase correction).
Then, we have the \emph{index} of the user, which is a sequence of QPSK symbols.
For example, this index can be a simple QPSK symbol ($\pm1\pm1j$), or a sequence of QPSK symols if the PoC has to consider more users.
With $\ell$ QPSK symbols, we can indeed fit at most $4^{\ell}/2 = 2^{2\ell-1}$ devices, and not $4^{\ell}$ due to the use of a conjugate index to send back acknowledgements from the base station.
Once the base station receives an uplink packet, it detects this index and transmits an acknowledgement which has the same frame structure, but where the index is the conjugate of the index of the uplink packet ($z \mapsto \overline{z}$, \emph{e.g.}, $1+j \mapsto 1-j$).
Thanks to this index, we can have several devices communicating with the same base station.
%
In turn, the end-device that receives the acknowledgement demodulates it, and checks if the index is the conjugate of its own index.
In this case, the \emph{Ack} was sent for him, and it knows that its packet has been received and decoded correctly by the base station.

@@

\begin{figure}[!t]
    \centering
    \includegraphics[width=0.70\linewidth]{SCEE_TestBed1.jpg}
    \vspace*{20pt}
    \includegraphics[width=0.70\linewidth]{SCEE_TestBed2.jpg}
    \caption{Two pictures showing the SCEE testbed \cite[Appendix~3]{Bodinier17}, taken in $2018$.}
    \label{fig:42:photosSCEETestBed}
\end{figure}

The boards have their own power supply, and are all connected to a local Ethernet switch, itself connected to a single laptop, running GNU/Linux and Ubuntu.
The pictures in Figure~\ref{fig:42:photosSCEETestBed} show the testbed used for these experiments.
% Octoclock
To ease the synchronization in both time and frequency between the boards representing the dynamic devices and the base station, we use an Octoclock \cite{OctoclockProduct}, also a product of Ettus Research,
% https://www.ettus.com/product/details/OctoClock-G
and coaxial cables connecting every platform to the Octoclock for time (PPS) and frequency synchronization, but this is not mandatory.

@@

    \item %[(2)]
    % \item[\textcolor{darkred}{\circled{2}}]
The second part is the interface of the intelligent device which is made of four parts.
At the top left, we observe the constellation of the transmitted packet \circled{a}.
At the bottom left, we have a time/frequency view of the lasts packets transmitted by the device \circled{b}.
We can see, in this view that the device transmitted its last $9$ packets in channels $\#3$ and $\#4$.
Then, at the top right of this interface \circled{c}, we can see the traffic observed by this device, where we have the interfering traffic (\textcolor{darkgreen}{green}), the uplink packets transmitted by this device (\textcolor{darkred}{red}) and the acknowledgements sent by the base station (\textcolor{darkblue}{blue}).
Colors in the ``waterfall'' represent the RF power level received at the device antenna. Hot colors are for closer elements, as for instance the device Tx antenna (reception) is close to the device Rx antenna (transmission), and consequently for the device waterfall these signals are colored in \textcolor{darkred}{red} (see graph \circled{c} in part \textcolor{darkred}{\boldcircled{2}}).
Finally, at the bottom right \circled{d}, we have four histograms showing the performance indicators of the chosen MAB algorithm (number of transmissions, number of successful transmissions, UCB indexes and success rates, in each channel).

    \item %[(3)]
    % \item[\textcolor{darkred}{\circled{3}}]
The last part is the interface of the base station, where we can see the traffic observed by the base station \circled{a} and the channels in which the last acknowledgements have been sent \circled{b}.
The observed traffic is coherent with \circled{c} of part \textcolor{darkred}{\boldcircled{2}}, as the elements are very close the one from the others on the testbed. Colors may change, as they depend on the exact distance between the different transmitters and receivers. See more details in \cite{MoyBesson2019}.
\end{enumerate}

@@

Similar results are obtained for overlapping channels, when dynamic devices are learning in the presence of multiple devices, all using the same learning algorithm.
However, our experimental testbed cannot run hundreds of intelligent devices.
Empirical results confirm the simulations presented in Section~\ref{sec:4:firstModel} (see Figure~\ref{fig:41:perf_learning}).
Such results are very encouraging, and illustrate well the various strong possibilities of MAB learning applied to IoT networks.

@@

\paragraph{Video.}
%
As depicted in Figure~\ref{fig:42:screenshotDemoYouTube} below,
we realized a $6$-minute \textbf{video} to sum-up our demonstration and advertise our work online, and it is available on the YouTube hosting platform, at \texttt{\href{https://youtu.be/HospLNQhcMk}{youtu.be/HospLNQhcMk}}.
The video shows examples of $3$ dynamic devices learning simultaneously, confirming the results of Figure~\ref{fig:42:plot_datafile_append_Uniform_vs_UCB_vs_TS} for overlapping channels.
It also shows the connections between the USRP boards, the Octoclock, the master laptop etc, completing the presentation of the SCEE testbed already shown in Figure~\ref{fig:42:photosSCEETestBed}.
% \textcolor{white}{Of course, the video was not seen much, because our work is useless! Yay!}

>>>>

We describe the way we implemented a demo where we evaluate MAB algorithms, used in combination with a pure ALOHA-based protocol, such as the ones employed in LPWAN.
This demonstration is the first implementation which aims at assessing the potential gain of MAB learning algorithms in IoT scenarios.
%
We use a test-bed designed in 2017 by the SCEE team at the Rennes campus of CentraleSupélec \cite[Appendix~3]{Bodinier17}, containing different USRP boards \cite{USRPDocumentation}, controlled by a single laptop running the GNU Radio software \cite{GNURadioDocumentation},
and where the intelligence of each device corresponds to a learning algorithm, implemented as a GNU Radio block \cite{GNURadioCompanionDocumentation} and written in Python or \texttt{C++}.

@@

Regarding the physical layer, we consider a QPSK constellation (\emph{Quadrature Phase-Shift Keying}, \eg, see references in \cite{Bodinier17}). Moreover, we use simplified packets composed of two parts.
The first part is the \emph{preamble} which is used for the purpose of synchronization (phase correction).
Then, we have the \emph{index} of the user, which is a sequence of QPSK symbols.
For example, this index can be a simple QPSK symbol ($\pm1\pm1j$), or a sequence of QPSK symols if the PoC has to consider more users.
With $\ell$ QPSK symbols, we can indeed fit at most $4^{\ell}/2 = 2^{2\ell-1}$ devices, and not $4^{\ell}$ due to the use of a conjugate index to send back acknowledgements from the base station.
Once the base station receives an up-link packet, it detects this index and transmits an acknowledgement which has the same frame structure, but where the index is the conjugate of the index of the up-link packet ($z \mapsto \overline{z}$, \emph{e.g.}, $1+j \mapsto 1-j$).
Thanks to this index, we can have several devices communicating with the same base station.
%
In turn, the end-device that receives the acknowledgement demodulates it, and checks if the index is the conjugate of its own index.
In this case, the \emph{Ack} was sent for him, and it knows that its packet has been received and decoded correctly by the base station.

@@

\begin{figure}[!t]
    \centering
    \includegraphics[width=0.70\linewidth]{SCEE_TestBed1.jpg}
    \vspace*{20pt}
    \includegraphics[width=0.70\linewidth]{SCEE_TestBed2.jpg}
    \caption{Two pictures showing the SCEE test-bed \cite[Appendix~3]{Bodinier17}, taken in $2018$.}
    \label{fig:42:photosSCEETestBed}
\end{figure}

The boards have their own power supply, and are all connected to a local Ethernet switch, itself connected to a single laptop, running GNU/Linux and Ubuntu.
The pictures in Figure~\ref{fig:42:photosSCEETestBed} show the test-bed used for these experiments.
% Octoclock
To ease the synchronization in both time and frequency between the boards representing the dynamic devices and the base station, we use an Octoclock \cite{OctoclockProduct}, also a product of Ettus Research,
% https://www.ettus.com/product/details/OctoClock-G
and coaxial cables connecting every platform to the Octoclock for time (PPS) and frequency synchronization, but this is not mandatory.

@@

    \item %[(2)]
    % \item[\textcolor{darkred}{\circled{2}}]
The second part is the interface of the intelligent device which is made of four parts.
At the top left, we observe the constellation of the transmitted packet \circled{a}.
At the bottom left, we have a time/frequency view of the lasts packets transmitted by the device \circled{b}.
We can see, in this view that the device transmitted its last $9$ packets in channels $\#3$ and $\#4$.
Then, at the top right of this interface \circled{c}, we can see the traffic observed by this device, where we have the interfering traffic (\textcolor{darkgreen}{green}), the up-link packets transmitted by this device (\textcolor{darkred}{red}) and the acknowledgements sent by the base station (\textcolor{darkblue}{blue}).
Colors in the ``waterfall'' represent the RF power level received at the device antenna. Hot colors are for closer elements, as for instance the device Tx antenna (reception) is close to the device Rx antenna (transmission), and consequently for the device waterfall these signals are colored in \textcolor{darkred}{red} (see graph \circled{c} in part \textcolor{darkred}{\boldcircled{2}}).
Finally, at the bottom right \circled{d}, we have four histograms showing the performance indicators of the chosen MAB algorithm (number of transmissions, number of successful transmissions, UCB indexes and success rates, in each channel).

    \item %[(3)]
    % \item[\textcolor{darkred}{\circled{3}}]
The last part is the interface of the base station, where we can see the traffic observed by the base station \circled{a} and the channels in which the last acknowledgements have been sent \circled{b}.
The observed traffic is coherent with \circled{c} of part \textcolor{darkred}{\boldcircled{2}}, as the elements are very close the one from the others on the test-bed. Colors may change, as they depend on the exact distance between the different transmitters and receivers. See more details in \cite{MoyBesson2019}.
\end{enumerate}

@@

Similar results are obtained for overlapping channels, when dynamic devices are learning in the presence of multiple devices, all using the same learning algorithm.
However, our experimental test-bed cannot run hundreds of intelligent devices.
Empirical results confirm the simulations presented in Section~\ref{sec:4:firstModel} (see Figure~\ref{fig:41:perf_learning}).
Such results are very encouraging, and illustrate well the various strong possibilities of MAB learning applied to IoT networks.

@@

\paragraph{Video.}
%
As depicted in Figure~\ref{fig:42:screenshotDemoYouTube} below,
we realized a $6$-minute \textbf{video} to sum-up our demonstration and advertise our work online, and it is available on the YouTube hosting platform, at \texttt{\href{https://youtu.be/HospLNQhcMk}{youtu.be/HospLNQhcMk}}.
The video shows examples of $3$ dynamic devices learning simultaneously, confirming the results of Figure~\ref{fig:42:plot_datafile_append_Uniform_vs_UCB_vs_TS} for overlapping channels.
It also shows the connections between the USRP boards, the Octoclock, the master laptop etc, completing the presentation of the SCEE test-bed already shown in Figure~\ref{fig:42:photosSCEETestBed}.
% \textcolor{white}{Of course, the video was not seen much, because our work is useless! Yay!}

### 2019-07-09 2-Chapters/4-Chapter/IEEE_WCNC_2019__DemoICT.git/main_appendix.tex
J'utilise moins "our ..."
Corrections fautes orthographes

The code corresponding to the following components is written in \texttt{C++} and \texttt{Python} for the hand-written blocks, and using the GUI of the GNU Radio Companion software.
The flow-graphs are saved as XML files, and the complete code of our demo is fully available online, open-sourced under GPLv3 license, at
\href{https://bitbucket.org/scee_ietr/malin-multi-arm-bandit-learning-for-iot-networks-with-grc}{\texttt{bitbucket.org/scee\_ietr/malin-multi-arm-bandit}\\\texttt{-learning-for-iot-networks-with-grc/}}.

@@

We note the difference between the base station and the end-devices.
The base station uses its Rx antenna to scan the entire range of the $K$ channels, at all time steps ; while dynamic devices use their Rx antennas only to listen for the acknowledgement sent by the base station if it received and decoded its uplink packet.
After each transmission, the dynamic device listens to one channel during a certain time interval, in the channel used for uplink transmission.

>>>>

The code corresponding to the following components is written in \texttt{C++} and \texttt{Python} for the hand-written blocks, and using the GUI of the GNU Radio Companion software.
The flow-graphs are saved as XML files, and the complete code of our proof-of-concept is fully available online, open-sourced under GPLv3 license, at
\href{https://bitbucket.org/scee_ietr/malin-multi-arm-bandit-learning-for-iot-networks-with-grc}{\texttt{bitbucket.org/scee\_ietr/malin-multi-arm-bandit}\\\texttt{-learning-for-iot-networks-with-grc/}}.

@@

We note the difference between the base station and the end-devices.
The base station uses its Rx antenna to scan the entire range of the $K$ channels, at all time steps ; while dynamic devices use their Rx antennas only to listen for the acknowledgement sent by the base station if it received and decoded its up-link packet.
After each transmission, the dynamic device listens to one channel during a certain time interval, in the channel used for up-link transmission.

### 2019-07-09 2-Chapters/4-Chapter/IEEE_WCNC__2019__Paper__BMBM.git/IEEE_WCNC__2019__Paper__BMBM.tex
Corrections fautes orthographes

Like in the previous model presented in Section~\ref{sec:4:firstModel},
regardless the type of devices, each of them follows a slotted ALOHA protocol \cite{Roberts75}, and has a probability $p>0$ to transmit a packet in a time slot.
We make the hypothesis that the transmission is successful if the channel is available, otherwise it fails upon radio collision.
The novelty compare to the previous model is that
in case of RF (uplink or downlink) collision that prevent the device to receive the \emph{Ack} from the gateway,
these devices will attempt to retransmit their packet up-to $\mathrm{MaxBackOff}$ times\footnote{~We denote it $\mathrm{MaxBackOff}$ instead of $M$ like we did in our paper \cite{Bonnefoi2019WCNC}, as $M$ is used in next Chapter~\ref{chapter:5}.},
with $\mathrm{MaxBackOff} \in\mathbb{N}^*$.
It is important to note that, every retransmission is carried out after a random back-off time, uniformly distributed in $\{0, \dots, m-1\}$, where $m \in\mathbb{N}^*$ is the length of the back-off interval
(note the difference between $m$ and $\mathrm{MaxBackOff}$).

@@

\begin{leftbar}[warningbar]  % XXX leftbar warningbar, comment if needed
	% \begin{small}  % XXX WARNING
		\textbf{\textcolor{red}{Small warning about notations: local (device) time vs global (gateway) time}.}
		%
		In the algorithms presented in this chapter, the time step $t$ does not refer to the \emph{global} time step (as seen from the gateway), but rather the current number of uplink transmissions (or retransmissions) that was carried out by the device.
		Each device wakes up whenever it has to send some data, following its own emission process (we restrict to a Bernoulli process of small probability, \eg, $p=10^{-3}$), and whenever it wakes up, it sends its packet, and will send again for at most $\mathrm{MaxBackOff}$ times (\eg, $\mathrm{MaxBackOff}=10$) until it receives an \Ack.
		%
		In other words, the time steps $t$ in the following algorithms denote the number of uplink packets sent by the device, and these uplink transmissions or retransmissions can happen at different (global or real-world) times.
		The local index $t$ is independent on the global time (\ie, the time of the gateway), and IoT device does not need to care about this difference.
		From the device point-of-view, the learning occurs at successive times (whenever the device wakes up), regardless of the global time.
	% \end{small}  % XXX WARNING
\end{leftbar}  % XXX leftbar warningbar, comment if needed

>>>>

Like in the previous model presented in Section~\ref{sec:4:firstModel},
regardless the type of devices, each of them follows a slotted ALOHA protocol \cite{Roberts75}, and has a probability $p>0$ to transmit a packet in a time slot.
We make the hypothesis that the transmission is successful if the channel is available, otherwise it fails upon radio collision.
The novelty compare to the previous model is that
in case of RF (up-link or down-link) collision that prevent the device to receive the \emph{Ack} from the gateway,
these devices will attempt to retransmit their packet up-to $\mathrm{MaxBackOff}$ times\footnote{~We denote it $\mathrm{MaxBackOff}$ instead of $M$ like we did in our paper \cite{Bonnefoi2019WCNC}, as $M$ is used in next Chapter~\ref{chapter:5}.},
with $\mathrm{MaxBackOff} \in\mathbb{N}^*$.
It is important to note that, every retransmission is carried out after a random back-off time, uniformly distributed in $\{0, \dots, m-1\}$, where $m \in\mathbb{N}^*$ is the length of the back-off interval
(note the difference between $m$ and $\mathrm{MaxBackOff}$).

@@

\begin{leftbar}[warningbar]  % XXX leftbar warningbar, comment if needed
	% \begin{small}  % XXX WARNING
		\textbf{\textcolor{red}{Small warning about notations: local (device) time vs global (gateway) time}.}
		%
		In the algorithms presented in this chapter, the time step $t$ does not refer to the \emph{global} time step (as seen from the gateway), but rather the current number of up-link transmissions (or retransmissions) that was carried out by the device.
		Each device wakes up whenever it has to send some data, following its own emission process (we restrict to a Bernoulli process of small probability, \eg, $p=10^{-3}$), and whenever it wakes up, it sends its packet, and will send again for at most $\mathrm{MaxBackOff}$ times (\eg, $\mathrm{MaxBackOff}=10$) until it receives an \Ack.
		%
		In other words, the time steps $t$ in the following algorithms denote the number of up-link packets sent by the device, and these up-link transmissions or retransmissions can happen at different (global or real-world) times.
		The local index $t$ is independent on the global time (\ie, the time of the gateway), and IoT device does not need to care about this difference.
		From the device point-of-view, the learning occurs at successive times (whenever the device wakes up), regardless of the global time.
	% \end{small}  % XXX WARNING
\end{leftbar}  % XXX leftbar warningbar, comment if needed
